Synthesizing Quotient Lenses

### 2018-02-23 papers/QuotientLenses/main.tex
edits

A {\em Quotient Regular Expression} (QRE) defines an external regular
expression $R$, an equivalence relation $\sim$ on the language
described by $R$, and a canonical set of representatives of $\sim$,
all at once.  QREs are constructed from a simple set of primitives
and composed using familiar regular operators
(choice `\cd{|}', concatenation `\cd{.}', sequential composition ``;''
and Kleene star `$^*$').

For example, to begin construction of our \bibtex{} format, we might
first define a whitespace format, which, externally, matches
any non-zero number of whitespace characters, and converts
any such whitespace into a single space character (its canonical form).
Such a definition uses the \kw{project} primitive to implement the
specification:

@@

While a projection chooses a single canonical representative for all elements
of a regular language, the \kw{squash} combinator maps data from one
regular language to another via a user-defined
function.  For instance,
if \codefont{comma\_name} describes the format for ``Conway, J. H.''
and \codefont{space\_name} describes the format for ``J. H. Conway''
and \codefont{s\_to\_c} is a function from one to the other then
the following instance of squash creates the desired canonizer.

>>>>

A {\em Quotient Regular Expression} (QRE) defines an external regular
expression $R$, an equivalence relation $\sim$ on the language
described by $R$, and a canonical set of representatives of $\sim$,
all at once.  QREs are constructed from a simple set of primitives
and composed using familiar regular operators
(choice `\cd{|}', concatenation `\cd{.}',
and Kleene star `$^*$').  Users may also treat QREs as pairs of
get and put functions and compose such pairs (the functional
composition operator `;').

For example, to begin construction of our \bibtex{} format, we might
first define a whitespace format, which, externally, matches
any non-zero number of whitespace characters, and converts
any such whitespace into a single space character (its canonical form).
Such a definition uses the \kw{project} primitive.

@@

Sometimes there are multiple disjoint representations of the same data.
In such situations, the \kw{squash} combinator creates a QRE that
allows external data to be in either format, and converts any
data in one format to the other.
For instance, assume
the \codefont{comma\_name} format describes ``Conway, J. H.''
and the \codefont{space\_name} format describes ``J. H. Conway''
and \codefont{s\_to\_c} is a function from the first to the second.  In this case,
the following instance of \kw{squash} creates the desired canonizer.
